---
title: "class12"
author: Ani A16647613
format: pdf
---

## Section 1: Proporation of G/G in a population

Downloaded a CSV file from the Ensemble from the lab directions

Now we will read the csv file

```{r}
mxl <- read.csv("373531-SampleGenotypes-Homo_sapiens_Variation_Sample_rs8067378 (1).csv")

head(mxl)
```

```{r}
table(mxl$Genotype..forward.strand.)
```

```{r}
table(mxl$Genotype..forward.strand.) / nrow(mxl) * 100
```
14.06% G/G

Now we are gonna look at population GBR (Great Britain)

```{r}
gbr <- read.csv("gbr.csv")
```

Find proportion of G/G
```{r}
round(table(gbr$Genotype..forward.strand.) / nrow(gbr) *100, 2)
```
This varient of childhood asthma is more frequent in GBR than MXL population, so noew we can look into this further:

##Section 4: Population Scale Analysis

Answering the following questions: 

Q13. Read the file into R and determine the sample size for each genotype and their corresponding median expression levels for each of these genotypes.


Q14. Generate a boxplot with a box per genotype, what would you inger from the relative expression value between A/A and G/G displayed in this plot? Does the SNP effect the expression of ORMDL3? 
How many samples do we have?

```{r}
expr <- read.table("project12part2.txt")
head(expr)
```
From the table function below, we can see the distribution amongst the different genotypes to see their sample sizes. 

```{r}
nrow(expr)
```
```{r}
table(expr$geno)
```
```{r}
library(ggplot2)
```

Make a boxggpplot 

```{r}
ggplot(expr) +
  aes(x= geno,y= exp, fill=geno) +
  geom_boxplot(notch= TRUE)
```

Above, we can interpret that having a G/G shows reduced expression of  the gene. We can also interpret that A/A shows high expression of the gene.
